Structuring College Access

### 2025-03-24 index.qmd
add rq2 tables

  group_headings <- unique(table_df$variable)
  nrows <- nrow(table_df) / length(group_headings)
  i <- 1

  k <- kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Mean', 'SD', 'P25', 'P50', 'P75'), 3)))

  for (g in group_headings) {
    k <- k %>% pack_rows(g, i, i + nrows - 1, colnum = 1, label_row_css = '')
    i <- i + nrows
  }

  k %>% 
    add_header_above(header = c(' ' = 1, '1980' = 5, '2000' = 5, '2020' = 5))
}

@@

@@

@@

@@

      writeLines(str_c(
        '<iframe data-src="./results/maps/rq2_map_',
        m,
        '.html" src="about:blank" width=100% height=100% style="max-height: 75%" allowtransparency="true"></iframe>'
      ))
      writeLines('<p class="fullscreen">Fullscreen</p>')
      notes_txt <- readLines(file.path(graphs_dir, rq, str_c('rq2_', m, '_race_row_plot_note.txt')), warn = FALSE)
      writeLines(
        str_replace(
          str_c('<div class="footnote">', paste0(notes_txt, collapse = "</br>"), '</div>'),
          'Excludes students with missing values for race\\. ', ''
        )
      )

@@

    # Build figure name and call your "no-heading" figure function
    graph_type <- paste0(g, "_", rc, "_plot_", native_order)
    insert_figure_aian(
      rq         = "rq2",
      metro      = m,
      graph_type = graph_type
    )
  }

>>>>

  headings <- unique(table_df$variable)
  headings_group <- rep(nrow(table_df) / length(headings), length(headings))
  names(headings_group) <- headings

  kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Mean', 'SD', 'P25', 'P50', 'P75'), 3))) %>% 
    pack_rows(index = headings_group, colnum = 1, label_row_css = '') %>% 
    add_header_above(header = c(' ' = 1, '1980' = 5, '2000' = 5, '2020' = 5))
}

insert_rq2_table <- function(metro, table_type, rq_name = 'rq2') {
  table_df <- readRDS(file.path(tables_dir, str_c(rq_name, '_table_', metro, '.rds')))[[table_type]]
  table_df[is.na(table_df)] <- NA_real_
  table_df$order_ids <- as_factor(table_df$order_ids)

  orders <- str_c(if_else(str_detect(unique(table_df$order_ids), '_'), 'Orders ', 'Order '), sub('_', ' & ', unique(table_df$order_ids)), ', ', unique(table_df$test_range))
  orders_group <- table(table_df$order_ids)
  names(orders_group) <- orders

  if (table_type == 'race') {
    tables_group <- rep(nrow(table_df) / length(orders) / length(unique(table_df$table)), length(unique(table_df$table)))
    names(tables_group) <- c('Count of purchased students', 'Purchased students by geomarket', 'Purchased students by racial/ethnic group')

    return(
      kbl(
        table_df %>%
          mutate(
            across(all, \(x) format(round(x, 0), big.mark = ',')),
            across(race_known, ~ if_else(table == 'col_pct_table' & eps_codename != 'All', sprintf('%.1f', .), format(round(., 0), big.mark = ','))),
            across(white:nhpi, ~ if_else(table == 'count_table' | (table == 'col_pct_table' & eps_codename == 'All'), format(round(., 0), big.mark = ','), sprintf('%.1f', .)))
          ) %>%
          select(-table, -order_ids, -test_range),
        align = 'lrrrrrrrrr', escape = F,
        col.names = c('', 'All', 'Race known', 'White', 'Asian', 'Black', 'Hispanic', '2+ races', 'AIAN', 'NHPI')
      ) %>% 
        pack_rows(index = orders_group, colnum = 1, label_row_css = 'background-color: #eaeaea; padding: 10px 7.5px;') %>% 
        pack_rows(index = rep(tables_group, length(orders)), colnum = 1, label_row_css = '')
    )
  } else if (table_type == 'firstgen') {
    tables_group <- rep(nrow(table_df) / length(orders) / length(unique(table_df$table)), length(unique(table_df$table)))
    names(tables_group) <- c('Count of purchased students', 'Purchased students by geomarket', 'Purchased students by first-generation status')

    return(
      kbl(
        table_df %>%
          mutate(
            across(all, \(x) format(round(x, 0), big.mark = ',')),
            across(known, ~ if_else(table == 'col_pct_table' & eps_codename != 'All', sprintf('%.1f', .), format(round(., 0), big.mark = ','))),
            across(no_col:not_first, ~ if_else(table == 'count_table' | (table == 'col_pct_table' & eps_codename == 'All'), format(round(., 0), big.mark = ','), sprintf('%.1f', .)))
          ) %>%
          select(-unknown, -table, -order_ids, -test_range),
        align = 'lrrrrr', escape = F,
        col.names = c('', 'All', 'First-gen known', 'No college', 'Some college', 'Not first-gen')
      ) %>% 
        pack_rows(index = orders_group, colnum = 1, label_row_css = 'background-color: #eaeaea; padding: 10px 7.5px;') %>% 
        pack_rows(index = rep(tables_group, length(orders)), colnum = 1, label_row_css = '')
    )
  } else if (table_type == 'race_firstgen') {
    group_df <- table_df %>% group_by(order_ids, stu_race_cb) %>% summarise(n = n())
    race_group <- group_df$n
    names(race_group) <- str_replace(str_to_title(group_df$stu_race_cb), 'Or', 'or')

    return(
      kbl(
        table_df %>% 
          mutate(
            across(all, \(x) format(round(x, 0), big.mark = ',')),
            across(row_no_col:row_not_first_gen, \(x) sprintf('%.1f', x))
          ) %>% 
          select(-stu_race_cb, -order_ids, -test_range),
        align = 'lrrrr', escape = F,
        col.names = c('', 'All', 'No college', 'Some college', 'Not first-gen')
      ) %>% 
        pack_rows(index = orders_group, colnum = 1, label_row_css = 'background-color: #eaeaea; padding: 10px 7.5px;') %>% 
        pack_rows(index = race_group, colnum = 1, label_row_css = '')
    )
  }
}

@@

      get_notes <- function(missing_txt = '') {
        str_replace(
          str_c('<div class="footnote">', paste0(readLines(file.path(graphs_dir, rq, str_c('rq2_', m, '_race_row_plot_note.txt')), warn = FALSE), collapse = "</br>"), '</div>'),
          'Excludes students with missing values for race\\. ', missing_txt
        )
      }

@@

        # Add RQ2 table
        writeLines(str_c("### Research question ", rq_num, " {.hide}"))
        writeLines(if_else(
          g == 'race',
          str_c('#### Racial/ethnic composition of purchased student profiles by Geomarket, ', heading_text),
          str_c('#### First-generation status of purchased student profiles by Geomarket, ', heading_text)
        ))

        writeLines('<div class="table-wrapper freeze-pane">')
        writeLines(insert_rq2_table(m, g))
        writeLines('</div>')
        writeLines(get_notes())

@@

      # Add RQ2B table
      writeLines(str_c("### Research question ", rq_num, " {.hide}"))
      writeLines(str_c("#### First-generation status by race for ", heading_text, " Geomarkets"))

      writeLines('<div class="table-wrapper freeze-pane">')
      writeLines(insert_rq2_table(m, 'race_firstgen'))
      writeLines('</div>')
      writeLines(get_notes('Excludes students who have missing values for race or parental education. '))

@@

      writeLines(str_c(
        '<iframe data-src="./results/maps/rq2_map_',
        m,
        '.html" src="about:blank" width=100% height=100% style="max-height: 75%" allowtransparency="true"></iframe>'
      ))
      writeLines('<p class="fullscreen">Fullscreen</p>')
      writeLines(get_notes())

@@

    # Build figure name and call your "no-heading" figure function
    graph_type <- paste0(g, "_", rc, "_plot_", native_order)
    insert_figure_aian(
      rq         = "rq2",
      metro      = m,
      graph_type = graph_type
    )

    # Add RQ2 table
    if ((g == 'firstgen' && rc == 'col') | (g == 'race' && rc == 'row')) {
      writeLines(paste0("### ", heading_text, " (AI/AN) {.hide}\n"))
      writeLines(if_else(
        g == 'race',
        str_c('#### Racial/ethnic composition of purchased student profiles by Geomarket, ', heading_text),
        str_c('#### First-generation status of purchased student profiles by Geomarket, ', heading_text)
      ))

      writeLines('<div class="table-wrapper freeze-pane">')
      writeLines(insert_rq2_table(m, g, 'rq2_aian'))
      writeLines('</div>')
      writeLines('<div class="footnote">Figure Notes:<br>- Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+.</div>')
    }
  }
